---
title: "HousingMarket"
author: "CGR"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
Sys.setlocale("LC_ALL", "English_United States.1252")

```


## Load libraries

```{r, include=FALSE}
library(lubridate)
library(ggpubr)
library(DataExplorer)
library(readr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(corrplot)
library(tidyverse)
library(stringr)
library(GGally)
library(gridExtra)
library(reshape)
library(groupdata2)
library(skimr)
library(patchwork)
library(reshape)
library(sjPlot)
library(jtools)
library(zoo)
library(data.table)
library(car)
library(tseries)
library(forecast)
library(MASS)
```

## 0. Import

```{r}
housesales <- read.csv("kaupskra.csv", header=T, fill=T, dec=",", sep=";", fileEncoding = 'UTF-8-BOM')

```


## 0. Clean
'postalcode','date',  'sellingprice', 'appraisal', 'age', 'msq', 'type', 'outcome'

```{r}
Housesales <- housesales %>%
  na.omit() %>%
  mutate(
    einflm = as.numeric(einflm) * 0.1,
    kaupverd = as.numeric(kaupverd) * 1000,
    fasteignamat = as.numeric(fasteignamat) * 1000,
    byggar = as.numeric(byggar),
    utgdag = ymd_hms(utgdag)
  ) %>%
  filter(byggar != 0) %>%
  filter(einflm > 25 & einflm < 200) %>%
  unite(ID,
        fepilog,
        heimilisfang,
        postnr,
        einflm,
        byggar,
        utgdag,
        tegund,
        sep = "_") %>%
  group_by(ID) %>%
  summarise(
    kaupverd = mean(kaupverd, na.rm = TRUE),
    fasteignamat = mean(fasteignamat, na.rm = TRUE),
    Count = n()
  ) %>%
  separate(
    ID,
    into = c(
      "fepilog",
      "heimilisfang",
      "postnr",
      "einflm",
      "byggar",
      "utgdag",
      "tegund"
    ),
    sep = "_"
  ) %>%
  mutate(outcome = (kaupverd - fasteignamat) / kaupverd) %>%
  group_by(utgdag) %>%
  mutate(inventory = n())


Housesales$einflm <- as.numeric(Housesales$einflm)
Housesales$byggar <- as.numeric(Housesales$byggar)
Housesales$kaupverd <- as.numeric(Housesales$kaupverd)

Housesales




```

#### Missingness
```{r}
Housesales %>%
  plot_missing()
```


#### Scatterplots
Assumptions for a linear model 
Linearity: All of the relationships between the independent variables and the dependent variable are not linear, so the change in the dependent variable is not proportional to the change in each independent variable.
```{r}
Housesales %>% 
  ggplot(aes(x = einflm, y = outcome)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.001) +
  theme_bw() 

Housesales %>% 
  ggplot(aes(x = byggar, y = outcome)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.001) +
  theme_bw() 

Housesales %>% 
  ggplot(aes(x = kaupverd, y = outcome)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.001) +
  theme_bw() 

Housesales %>% 
  ggplot(aes(x = fasteignamat, y = outcome)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.001) +
  theme_bw() 
```

#### Outlier Analysis

Look for outliers in the original data before taking the means and medians below. 
```{r}
boxplot(Housesales$einflm)
boxplot(Housesales$byggar)
boxplot(Housesales$kaupverd)
boxplot(Housesales$fasteignamat)
```

##### Detailed Outlier Analysis

```{r}
#Housesales$outcome_p = Housesales$outcome - min(Housesales$outcome) +100
Housesales$nuvirdi = abs(Housesales$outcome)  #because the lm function does not work on zero values
fit_realscale <- lm(nuvirdi ~ postnr + einflm + byggar + inventory, Housesales)
#boxcox(fit_realscale, plotit=T, lambda = seq(-0.05, 0.25, len = 30))
#lambda_opt <- fit_realscale$lambda
```

```{r}
diag_realscale <- fortify(fit_realscale)
preal <- ggplot(diag_realscale, aes(x=.fitted,y=nuvirdi))+geom_point()
preal <- preal + xlab("Fitted") + ylab("Nuvirdi") + labs(title = "Nuvirdi vs. fitted values")
preal
```

```{r}
Housesales$index2 <- c(1:length(Housesales$nuvirdi))
set.seed(123)
train <- sample_frac(Housesales, 3/4)
test <- Housesales[-train$index2,]
train_index <- c(1:length(train$nuvirdi))
train$index <- train_index 
test_index <- c(1:length(test$nuvirdi))
test$index <- test_index 
fit <- lm(nuvirdi ~ postnr + einflm + byggar + inventory, Housesales)
```

```{r}
diag <- fortify(fit)
diag$.index = c(1:length(diag$.resid))
diag$.jack<-rstudent(fit)
```

##### Residuals
```{r}
res_sd <- summary(fit)$sigma
fit$sigma
```
```{r}
res_index <- ggplot(diag, aes(x = seq(1:length(.resid)), y = .resid)) + geom_point()
res_index <- res_index + geom_hline(yintercept = 0, col = "red", linetype = "dashed")
res_index <- res_index + geom_hline(yintercept = 2 * res_sd, col = "blue", linetype = "dashed")
res_index <- res_index + geom_hline(yintercept = -2 * res_sd, col = "blue",
linetype = "dashed")
res_index <- res_index + geom_hline(yintercept = 3 * res_sd, col = "green",
linetype = "dashed")
res_index <- res_index + geom_hline(yintercept = -3 * res_sd, col = "green",
linetype = "dashed")
res_index <- res_index + xlab("Index") + ylab("Residuals") + labs(title = "Index plot of residuals")
res_index <- res_index + geom_text(aes(label = ifelse(abs(.resid) > 2 * res_sd,
.index, "")), hjust = -0.5, vjust = 0.4)
res_index
which(abs(diag$.resid)>2 * res_sd)
```

##### Leverage
```{r}
p<-length(coef(fit))
n<-length(fitted(fit))
lev_index<-ggplot(diag, aes(x=seq(1:length(.hat)),y=.hat))+geom_point()
lev_index<-lev_index+geom_hline(yintercept=2*p/n, col="red", linetype="dashed")
lev_index<-lev_index+xlab("Index")+ylab("Leverages")
lev_index<-lev_index+geom_text(aes(label=ifelse(.hat>2*p/n,.index,"")),hjust=0, vjust=0)
lev_index
which(diag$.hat>2*p/n)
```

##### Studentized
```{r}
Stres_sd <- sd(diag$.stdresid, na.rm = TRUE)
p3 <- ggplot(diag,aes(x=seq(1:length(.stdresid)),y=.stdresid))+geom_point()
p3 <- p3 + geom_hline(yintercept=0, col="red", linetype="dashed")
p3 <- p3 + geom_hline(yintercept=2*Stres_sd, col="blue", linetype="dashed")
p3 <- p3 + geom_hline(yintercept=-2*Stres_sd, col="blue", linetype="dashed")
p3 <- p3 + geom_hline(yintercept=3*Stres_sd, col="green", linetype="dashed")
p3 <- p3 + geom_hline(yintercept=-3*Stres_sd, col="green", linetype="dashed")
p3 <- p3 + xlab("Index")+ylab("Studentized residuals")+labs(title = "Index plot of std.")
p3 <- p3 + geom_text(aes(label=ifelse(abs(.stdresid)>2*Stres_sd,.index,"")),hjust=-0.5, vjust=0.4)
p3
which(abs(diag$.stdresid)>2*Stres_sd)
```

##### qqplot studentized residuals
```{r}
df <- length(diag$.stdresid) - length(summary(fit)$coefficients[,1])
stdres <- diag$.stdresid[!is.na(diag$.stdresid)]
tqqplot(stdres,df)
```
##### Jacknife Residuals
```{r}
# Add jackknife residuals to diag dataframe
diag$.jack <- rstudent(fit) 
jackres <- diag$.jack[!is.na(diag$.jack)]
jackindex <- diag$.index[!is.na(diag$.jack)]


# Jackknife index plot
p4<-ggplot(diag, aes(x=seq(1:length(.jack)),y=.jack, na.rm = TRUE))+geom_point()
p4<-p4+geom_hline(yintercept=0, col="red", linetype="dashed")
p4<-p4+xlab("Index")+ylab("Jackknife residuals")
p4<-p4+geom_text(aes(label=ifelse(abs(.jack)>3,.index,"")),hjust=-0.1)
p4
which(diag$.jack>3)
```

##### qqplot jacknife residuals
```{r}
df <- length(jackres) - length(summary(fit)$coefficients[,1]) - 1
tqqplot(jackres,df)
```
If the model violates some of the assumptions of linear regression (such as normality or homoscedasticity), the Jackknife and studentized residuals may not always identify the same influential observations.


##### Bonferroni Correction & Outlier test

Test if points number no. 5344 6983 7085 7087 are outliers by computing the probability of having a larger absolute value than this observation. Degrees of freedom of the Jackknife residuals are n−1−p=13344-1-4= 13339
```{r}
prob5344 <- 2*(1 - pt(abs(diag$.jack[5344]),df))
prob6983 <- 2*(1 - pt(abs(diag$.jack[6983]),df))
prob7085 <- 2*(1 - pt(abs(diag$.jack[7085]),df))
prob7087 <- 2*(1 - pt(abs(diag$.jack[7087]),df))

alpha <- 0.05
bonf <- alpha/length(diag$.jack)
bonfprob <- c(prob5344,prob6983,prob7085,prob7087,bonf)
(bonfprob)

bonfprob_v <- bonfprob < bonf
print(bonfprob_v)

```
All 4 points have a p-value smaller than the Bonferroni corrected significance level, thus, they should be considered as outliers.




##### Cook's Distance

```{r}
p5 <- ggplot(diag, aes(x=seq(1:length(.cooksd)),y=.cooksd, na.rm = TRUE))+geom_point()
p5 <- p5 + xlab("Index")+ylab("Cooks distance") + theme(legend.position="none")
p5 <- p5 + geom_text(aes(label=ifelse(abs(.cooksd)>0.021,.index,"")))
p5

which(diag$.cooksd >0.021)
# 6983 7085 7087
```

Points 6983 7085 7087 are both Cook's Distance outliers and bonferroni corrected outliers. 

##### Closer Look at the outliers

```{r}
fit_cooks1 <- lm(nuvirdi ~ postnr + einflm + byggar + inventory, Housesales[-6983,])
relative_beta_cooks1 = abs(fit$coefficients - fit_cooks1$coefficients)/summary(fit)$coefficients[,2]

fit_cooks2 <- lm(nuvirdi ~ postnr + einflm + byggar + inventory, Housesales[-7085,])
relative_beta_cooks2 = abs(fit$coefficients - fit_cooks2$coefficients)/summary(fit)$coefficients[,2]

fit_cooks3 <- lm(nuvirdi ~ postnr + einflm + byggar + inventory, Housesales[-7087,])
relative_beta_cooks3 = abs(fit$coefficients - fit_cooks3$coefficients)/summary(fit)$coefficients[,2]

```


The absolute change in the parameter estimates with respect to the standard error of each parameter when points no. 6983 7085 7087 are removed. These points have substantial influence on several of the coefficients.
```{r}
((relative_beta_cooks1))
which(relative_beta_cooks1 > 1)
((relative_beta_cooks2))
which(relative_beta_cooks2 > 1)
((relative_beta_cooks3))
which(relative_beta_cooks3 > 1)
```

#incomplete
```{r}
predict_verd_train <- exp(predict(fit,train))
predict_verd_train_cooks1 <- exp(predict(fit_cooks1,train))
perc_pred_cooks1 = sort(100*abs(predict_verd_train - predict_verd_train_cooks1)/abs(predict_verd_train),decreasing=TRUE)
(perc_pred_cooks1[1:8])
```

## 1. Housesales_by_date

```{r}
Housesales_by_date <- Housesales %>%
  ungroup() %>%
  mutate(tegund = gsub(" ", "", tegund)) %>%
  filter(tegund %in% c("Einbýli", "Fjölbýli", "Sérbýli")) %>%
  mutate(einflm = as.numeric(einflm),
         byggar = as.numeric(byggar)) %>%
  group_by(utgdag) %>%
  summarise(
    byggar_mean = mean(byggar, na.rm = TRUE),
    byggar_median = median(byggar, na.rm = TRUE),
    einflm_mean = mean(einflm, na.rm = TRUE),
    einflm_median = median(einflm, na.rm = TRUE),
    outcome_mean = mean(outcome, na.rm = TRUE),
    inventory = n()
  ) %>% 
  mutate(across(where(is.numeric), round, 2))

Housesales_by_date

```

### Numerical data

```{r fig.height=11, fig.width=11}
Housesales_by_date %>% 
  plot_histogram()
```

### Categorical 

```{r}
Housesales_by_date %>% 
  plot_bar()
```


## 1. Housesales_by_date_pc

```{r}
Housesales_by_date_pc <- Housesales %>%
  ungroup() %>%
  mutate(tegund = gsub(" ", "", tegund)) %>%
  filter(tegund %in% c("Einbýli", "Fjölbýli", "Sérbýli")) %>%
  mutate(einflm = as.numeric(einflm),
         byggar = as.numeric(byggar)) %>%
  group_by(utgdag, postnr) %>%
  summarise(
    byggar_mean = mean(byggar, na.rm = TRUE),
    byggar_median = median(byggar, na.rm = TRUE),
    einflm_mean = mean(einflm, na.rm = TRUE),
    einflm_median = median(einflm, na.rm = TRUE),
    outcome_mean = mean(outcome, na.rm = TRUE),
    inventory = n()
  ) %>% 
  ungroup() %>% 
  mutate(across(where(is.numeric), round, 2))

PostNr <- Housesales_by_date_pc %>%
  count(postnr) %>% 
  mutate(postnr = as.numeric(postnr)) %>% 
  filter(between(postnr, 100, 299) | between(postnr, 600, 699)) %>% 
  filter(n > 1000) %>% 
  arrange(-n) %>% 
  pull(postnr)
  

Housesales_by_date_pc %>%
  filter(postnr %in% PostNr) %>% 
  distinct(postnr)

Housesales_by_date_pc

```

### Numerical data

```{r fig.height=11, fig.width=11}
Housesales_by_date_pc %>% 
  plot_histogram()
```

### Categorical 

```{r}
Housesales_by_date_pc %>% 
  plot_bar()
```


## 0. Clean: Rental Price, Rental Inventory, Meters Squared, and # of bedrooms

```{r}
rentalprice <-read.csv("leiguskra.csv", header=T, fill=T, dec=",", sep=",", fileEncoding = 'UTF-8-BOM')

#NAs are found in postnr & byggar
rentalprice %>% 
  plot_missing()

# convert date column to date format
rentalprice$UTGDAG <- as.Date(rentalprice$UTGDAG, format = "%Y-%m-%d %H:%M:%OS")

rentalprice_c <- rentalprice %>% 
  mutate(date = ymd(UTGDAG),
         STAERD = as.numeric(STAERD))%>%
  dplyr::select(-c(LEIGUSAMNINGUR, EMNR, SKJALNR, FASTNUM, HEINUM, HEIMILISFANG, SVFN, UTGDAG, SVEITARFELAG, THINGLYSTDAGS, DAGSFRA, DAGSTIL, FJ_HERBERGI, OTIMABUNDID, VISITALA, DAGS_VISITALA, ONOTHAEFUR_SAMNINGUR)) 

colnames(rentalprice_c) <- c('POSTALCODE', 'PRICE', 'SIZE', 'TYPE', 'date')

rentalprice_s <- subset(rentalprice_c, TYPE %in% c("Fjölbýli", "Einbýli", "Sérbýli"))

rentalprice_f <- filter(rentalprice_s, SIZE > 25 & SIZE < 200 )

summary(rentalprice_f$PRICE)
rentalprice_f <- filter(rentalprice_f, PRICE < 400000  )

head(rentalprice_f)
  
```

## 0.Clean: Interest
```{r}
interest <-read.csv("meginvextir_US_Date.csv", header=T, fill=T, dec=",", sep=";")

interest %>% 
  plot_missing()

interest_c <- interest %>% 
  mutate(date = ymd(date))

head(interest)
```

## full_data Datatable 

```{r}
full_data <- interest_c %>% 
  inner_join(housesales_f, by = "date") %>%
  inner_join(rentalprice_f, by = c("date")) %>%
  group_by(date) %>% 
  summarise(mean_interest_rate = mean(interest_rate), 
            mean_age = mean(age),
            mean_msq = mean(msq), 
            mean_outcome =mean(outcome), 
            mean_PRICE = mean(PRICE), 
            mean_SIZE = mean(SIZE) 
            )

fulldata_cp <- full_data %>%
  dplyr::select(-c(date, mean_outcome))

```

### Numerical data

```{r fig.height=11, fig.width=11}
full_data %>% 
  plot_histogram()
```

### Bootstrapping
```{r}
hist(housesales_g$msq)
tmp_mean <- replicate(10000, mean(sample(housesales_g$msq, replace=T), na.rm=T))
quantile(tmp_mean, 0.025)
quantile(tmp_mean, 0.975)
hist(tmp_mean)


tmp_med <- replicate(10000, mean(sample(housesales_g$outcome, replace=T), na.rm=T))
quantile(tmp_med, 0.025)
quantile(tmp_med, 0.975)
hist(tmp_med)
```


### Categorical 

```{r eval=FALSE, include=FALSE}
full_data %>% 
  plot_bar()
```

### Outlier Analysis
```{r}
boxplot(full_data$mean_interest_rate)
boxplot(full_data$mean_age)
boxplot(full_data$mean_msq)
boxplot(full_data$mean_outcome)
boxplot(full_data$mean_PRICE)
boxplot(full_data$mean_SIZE)
```

## Scatterplot & Transformations
```{r}
full_data$mean_outcomeT = full_data$mean_outcome - min(full_data$mean_outcome) +100
model <- lm(mean_outcomeT ~ . - date, full_data)
boxcox(model)

full_data %>% 
  ggplot(aes(x = mean_interest_rate, y = mean_outcomeT)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.001) +
  theme_bw() 

full_data %>% 
  ggplot(aes(x = mean_age, y = mean_outcomeT)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.001) +
  theme_bw() 

full_data %>% 
  ggplot(aes(x = mean_msq, y = mean_outcomeT)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.001) +
  theme_bw() 

full_data %>% 
  ggplot(aes(x = mean_PRICE, y = mean_outcomeT)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.001) +
  theme_bw() 

full_data %>% 
  ggplot(aes(x = mean_SIZE, y = mean_outcomeT)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.001) +
  theme_bw() 
```


```{r}

full_data$cubed <- (full_data$mean_outcome)^(1/3)
fit1 <- lm(cubed ~ mean_interest_rate + mean_age + mean_msq + mean_PRICE + mean_SIZE, data=full_data)
summary(fit1)
vif(fit1)
plot(fit1)
```

```{r}
full_data %>%
  plot_correlation()
```

```{r}
#without cubic root transformation
fit1 <- lm(mean_outcome ~ mean_interest_rate + mean_age + mean_msq + mean_PRICE + mean_SIZE, data=full_data)
summary(fit1)
vif(fit1)
plot(fit1)

full_data1$index2 <- c(1:length(full_data1$mean_msq))

set.seed(3689)
train <- sample_frac(full_data1, 3/4)
test <- full_data1[-train$index2,]

train_index <- c(1:length(train$mean_msq))
train$index <- train_index 
test_index <- c(1:length(test$mean_msq))
test$index <- test_index 

diag_realscale <- fortify(fit1)
preal <- ggplot(diag_realscale, aes(x=.fitted,y=mean_outcome))+geom_point()
preal <- preal + xlab("Fitted") + ylab("Nuvirdi") + labs(title = "Nuvirdi vs. fitted values")
preal
```

```{r}
full_data1 <- full_data %>% 
  dplyr::rename(nuvirdi = cubed)

full_data1$index2 <- c(1:length(full_data1$mean_msq))

set.seed(3689)
train <- sample_frac(full_data1, 3/4)
test <- full_data1[-train$index2,]

train_index <- c(1:length(train$mean_msq))
train$index <- train_index 
test_index <- c(1:length(test$mean_msq))
test$index <- test_index 

```

```{r}
fit_realscale <- lm(nuvirdi ~ .,data = train)
```

```{r}
diag_realscale <- fortify(fit_realscale)
preal <- ggplot(diag_realscale, aes(x=.fitted,y=nuvirdi))+geom_point()
preal <- preal + xlab("Fitted") + ylab("Nuvirdi") + labs(title = "Nuvirdi vs. fitted values")
preal
```

```{r}
cor(train[2:7])>0.9
```

```{r}
fit <- lm(nuvirdi ~ .,data = train)
diag <- fortify(fit)
diag$.index = c(1:length(diag$.resid))
diag$.jack<-rstudent(fit)

res_sd <- summary(fit)$sigma
fit$sigma
```
## Extreme points
```{r}
res_index <- ggplot(diag, aes(x = seq(1:length(.resid)), y = .resid)) + geom_point()
res_index <- res_index + geom_hline(yintercept = 0, col = "red", linetype = "dashed")
res_index <- res_index + geom_hline(yintercept = 2 * res_sd, col = "blue", linetype = "dashed")
res_index <- res_index + geom_hline(yintercept = -2 * res_sd, col = "blue",
linetype = "dashed") # 2 standard dev.
res_index <- res_index + geom_hline(yintercept = 3 * res_sd, col = "green",
linetype = "dashed") #3 std
res_index <- res_index + geom_hline(yintercept = -3 * res_sd, col = "green",
linetype = "dashed")
res_index <- res_index + xlab("Index") + ylab("Residuals") + labs(title = "Index plot of residuals")
res_index <- res_index + geom_text(aes(label = ifelse(abs(.resid) > 2 * res_sd,
.index, "")), hjust = -0.5, vjust = 0.4)
res_index

unique(ifelse(abs(diag$.resid)>2 * res_sd, diag$.index, ""))
which(abs(diag$.resid)>2 * res_sd)
```
## Leverage
```{r}
p<-length(coef(fit))
n<-length(fitted(fit))
lev_index<-ggplot(diag, aes(x=seq(1:length(.hat)),y=.hat))+geom_point()
lev_index<-lev_index+geom_hline(yintercept=2*p/n, col="red", linetype="dashed")
lev_index<-lev_index+xlab("Index")+ylab("Leverages")
lev_index<-lev_index+geom_text(aes(label=ifelse(.hat>2*p/n,.index,"")),hjust=0, vjust=0)
lev_index
which(diag$.hat>2*p/n)
```

## studentized 
```{r}
Stres_sd <- sd(diag$.stdresid, na.rm = TRUE)
p3 <- ggplot(diag,aes(x=seq(1:length(.stdresid)),y=.stdresid))+geom_point()
p3 <- p3 + geom_hline(yintercept=0, col="red", linetype="dashed")
p3 <- p3 + geom_hline(yintercept=2*Stres_sd, col="blue", linetype="dashed")
p3 <- p3 + geom_hline(yintercept=-2*Stres_sd, col="blue", linetype="dashed")
p3 <- p3 + geom_hline(yintercept=3*Stres_sd, col="green", linetype="dashed")
p3 <- p3 + geom_hline(yintercept=-3*Stres_sd, col="green", linetype="dashed")
p3 <- p3 + xlab("Index")+ylab("Studentized residuals")+labs(title = "Index plot of std.")
p3 <- p3 + geom_text(aes(label=ifelse(abs(.stdresid)>2*Stres_sd,.index,"")),hjust=-0.5, vjust=0.4)
p3
which(abs(diag$.stdresid)>2*Stres_sd)
```

```{r}
df <- length(diag$.stdresid) - length(summary(fit)$coefficients[,1])
stdres <- diag$.stdresid[!is.na(diag$.stdresid)]
tqqplot(stdres,df)

hist(stdres)
```


### timeseries 
```{r}

housingmarket <- read.zoo(full_data)
timser <- as.ts(housingmarket)
```

